CrypTool Script \\ Mathematics and Cryptography

### 2002-10-29 script/en/chapters/primes.tex
be fixes

The first $40$ values are prime numbers (which have the obvious regularity that
their difference starts with $2$ and increases by $2$ each time), but the $41$th
and $42$th values are not prime numbers.     It is easy to see that $f(41)$ hl
cannot be a prime number:
    $f(41) = 41^2 + 41 + 41 = 41 (41 + 1 + 1) = 41 \cdot 43$.
\pagebreak
  \item $f(n) = n^2 - 79 \cdot n + 1,601$: \\
    This function delivers prime numbers for all values from $n=0$ to $n=79$.
Unfortunately $f(80) = 1,681 = 11 \cdot 151$ is not a prime number. To this
date, no function has been found that delivers more prime numbers in a row. On
the other hand, each prime occurs twice (first in the decreasing then in the
increasing sequence), which means that the algorithm delivers a total of 40
difference prime values (the same ones as the function from point 10).
$$
\begin{array}{|ll||ll|}
\hline
f(0) = 1,601    & \mapsto {\rm ~prime} &  f(28) = 173    & \mapsto {\rm ~prime}
\\
f(1) = 1,523    & \mapsto {\rm ~prime} &  f(29) = 151    & \mapsto {\rm ~prime}
\\
f(2) = 1,447    & \mapsto {\rm ~prime} &  f(30) = 131 & \mapsto {\rm ~prime} \\
f(3) = 1,373    & \mapsto {\rm ~prime} &  f(31) = 113 & \mapsto {\rm ~prime} \\
f(4) = 1,301    & \mapsto {\rm ~prime} &  f(32) = 97 & \mapsto {\rm ~prime} \\
f(5) = 1,231    & \mapsto {\rm ~prime} &  f(33) = 83 & \mapsto {\rm ~prime} \\
f(6) = 1,163    & \mapsto {\rm ~prime} &  f(34) = 71 & \mapsto {\rm ~prime} \\
f(7) = 1,097    & \mapsto {\rm ~prime} &  f(35) = 61 & \mapsto {\rm ~prime} \\
f(8) = 1,033    & \mapsto {\rm ~prime} &  f(36) = 53 & \mapsto {\rm ~prime} \\
f(9) = 971  & \mapsto {\rm ~prime} &  f(37) = 47 & \mapsto {\rm ~prime} \\
f(10) = 911 & \mapsto {\rm ~prime} &  f(38) = 43 & \mapsto {\rm ~prime} \\
f(11) = 853 & \mapsto {\rm ~prime} &  f(39) = 41 & \mapsto {\rm ~prime} \\
f(12) = 797 & \mapsto {\rm ~prime} &             &                     \\
f(13) = 743 & \mapsto {\rm ~prime} &  f(40) = 41 & \mapsto {\rm ~prime} \\
f(14) = 691 & \mapsto {\rm ~prime} &  f(41) = 43 & \mapsto {\rm ~prime} \\
f(15) = 641 & \mapsto {\rm ~prime} &  f(42) = 47 & \mapsto {\rm ~prime} \\
f(16) = 593 & \mapsto {\rm ~prime} &  f(43) = 53 & \mapsto {\rm ~prime} \\
f(17) = 547 & \mapsto {\rm ~prime} &  \cdots  &  \\
f(18) = 503 & \mapsto {\rm ~prime} &  f(77) = 1,447  & \mapsto {\rm ~prime} \\
f(19) = 461 & \mapsto {\rm ~prime} &  f(78) = 1,523  & \mapsto {\rm ~prime} \\
f(20) = 421 & \mapsto {\rm ~prime} &  f(79) = 1,601  & \mapsto {\rm ~prime} \\
f(21) = 383 & \mapsto {\rm ~prime} &  f(80) = 11 \cdot 151 & \mapsto {\rm
~NOT~prime!} \\
f(22) = 347 & \mapsto {\rm ~prime} &  f(81) = 41 \cdot 43 & \mapsto {\rm
~NOT~prime!} \\
f(21) = 383 & \mapsto {\rm ~prime} &  f(82) = 1,847  & \mapsto {\rm ~prime} \\
f(22) = 347 & \mapsto {\rm ~prime} &  f(83) = 1,933  & \mapsto {\rm ~prime} \\
f(23) = 313 & \mapsto {\rm ~prime} &  f(84) = 43 \cdot 47 &  \mapsto {\rm
~NOT~prime!} \\
f(24) = 281 & \mapsto {\rm ~prime} & & \\
f(25) = 251 & \mapsto {\rm ~prime} & & \\
f(26) = 223 & \mapsto {\rm ~prime} & & \\
f(27) = 197 & \mapsto {\rm ~prime} & & \\
\hline
\end{array} 
$$
  \item Polynomial functions $f(x) = a_n x^n + a_{n-1}x^{n-1} + \cdots + a_1 x^1 +
a_0$  ($a_i$ in ${\mathbb Z}$, $n \geq 1$):

>>>>

The first $40$ values are prime numbers (which have the obvious regularity that
their difference starts with $2$ and increases by $2$ each time), but the $41$th
and $42$th values are not prime numbers.     It is easy to see that $f(41)$ hl
cannot be a prime number:
    $f(41) = 41^2 + 41 + 41 = 41 (41 + 1 + 1) = 41 \cdot 43$.
\pagebreak
  \item $f(n) = n^2 - 79 \cdot n + 1,601$: \\
    This function delivers prime numbers for all values from $n=0$ to $n=79$.
Unfortunately $f(80) = 1,681 = 11 \cdot 151$ is not a prime number. To this
date, no function has been found that delivers more prime numbers in a row. On
the other hand, each prime occurs twice (first in the decreasing then in the
increasing sequence), which means that the algorithm delivers a total of 40
difference prime values (the same ones as the function from point 10).
$$
\begin{array}{|ll||ll|}
\hline
f(0) = 1,601    & \mapsto {\rm ~prime} &  f(28) = 173    & \mapsto {\rm ~prime}
\\
f(1) = 1,523    & \mapsto {\rm ~prime} &  f(29) = 151    & \mapsto {\rm ~prime}
\\
f(2) = 1,447    & \mapsto {\rm ~prime} &  f(30) = 131 & \mapsto {\rm ~prime} \\
f(3) = 1,373    & \mapsto {\rm ~prime} &  f(31) = 113 & \mapsto {\rm ~prime} \\
f(4) = 1,301    & \mapsto {\rm ~prime} &  f(32) = 97 & \mapsto {\rm ~prime} \\
f(5) = 1,231    & \mapsto {\rm ~prime} &  f(33) = 83 & \mapsto {\rm ~prime} \\
f(6) = 1,163    & \mapsto {\rm ~prime} &  f(34) = 71 & \mapsto {\rm ~prime} \\
f(7) = 1,097    & \mapsto {\rm ~prime} &  f(35) = 61 & \mapsto {\rm ~prime} \\
f(8) = 1,033    & \mapsto {\rm ~prime} &  f(36) = 53 & \mapsto {\rm ~prime} \\
f(9) = 971  & \mapsto {\rm ~prime} &  f(37) = 47 & \mapsto {\rm ~prime} \\
f(10) = 911 & \mapsto {\rm ~prime} &  f(38) = 43 & \mapsto {\rm ~prime} \\
f(11) = 853 & \mapsto {\rm ~prime} &  f(39) = 41 & \mapsto {\rm ~prime} \\
f(12) = 797 & \mapsto {\rm ~prime} &  f(40) = 41 & \mapsto {\rm ~prime} \\               
f(13) = 743 & \mapsto {\rm ~prime} &  f(41) = 43 & \mapsto {\rm ~prime} \\               
f(14) = 691 & \mapsto {\rm ~prime} &  f(42) = 47 & \mapsto {\rm ~prime} \\               
f(15) = 641 & \mapsto {\rm ~prime} &  f(43) = 53 & \mapsto {\rm ~prime} \\               
f(16) = 593 & \mapsto {\rm ~prime} &  \cdots  &  \\                                      
f(17) = 547 & \mapsto {\rm ~prime} &  f(77) = 1,447  & \mapsto {\rm ~prime} \\           
f(18) = 503 & \mapsto {\rm ~prime} &  f(78) = 1,523  & \mapsto {\rm ~prime} \\           
f(19) = 461 & \mapsto {\rm ~prime} &  f(79) = 1,601  & \mapsto {\rm ~prime} \\           
f(20) = 421 & \mapsto {\rm ~prime} &  f(80) = 11 \cdot 151 & \mapsto {\rm ~NOT~prime!} \\
f(21) = 383 & \mapsto {\rm ~prime} &  f(81) = 41 \cdot 43 & \mapsto {\rm ~NOT~prime!} \\ 
f(22) = 347 & \mapsto {\rm ~prime} &  f(82) = 1,847  & \mapsto {\rm ~prime} \\           
f(21) = 383 & \mapsto {\rm ~prime} &  f(83) = 1,933  & \mapsto {\rm ~prime} \\           
f(22) = 347 & \mapsto {\rm ~prime} &  f(84) = 43 \cdot 47 &  \mapsto {\rm ~NOT~prime!} \\
f(23) = 313 & \mapsto {\rm ~prime} & & \\
f(24) = 281 & \mapsto {\rm ~prime} & & \\
f(25) = 251 & \mapsto {\rm ~prime} & & \\
f(26) = 223 & \mapsto {\rm ~prime} & & \\
f(27) = 197 & \mapsto {\rm ~prime} & & \\
\hline
\end{array} 
$$
  \item Polynomial functions $f(x) = a_n x^n + a_{n-1}x^{n-1} + \cdots + a_1 x^1 +
a_0$  ($a_i$ in ${\mathbb Z}$, $n \geq 1$):
